---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r library}

rm(list=ls())
# library(devtools)
# install_github("rOpenHealth/rEHR")
require(rpcdsearch)
library(rEHR)
library(readxl)
library(lubridate)
library(readr)
library(plyr)
library(dplyr)
library(reshape2)
library(xlsx)
library(knitr)

```

Incidence density matching: los controles se seleccionan para un caso particular en el momento de entrada en la intervención de éste; de modo que se permita que el mismo individuo pueda ser seleccionado como control para mas de un caso.. 

Según la función para crear la cohorte la fecha marcada en "diagnostic_start" es la que define la fecha de partida: the diagnosis_start argument is used to include the diagnosis date in the definition of the start dates for the patients. [de los controles deberia de usar de star date la primera de los casos? [y como end? el fallecimiento]] 


```{r table nav_ casos,controles}
rm(list=ls())


rec01<-colwise(function(x){
  return(ifelse(x >1, 1, x))})  # codifica >1=1
recn0<-colwise(function(x){
  return(ifelse(x <0, 0, x))})  # codifica <0=0
recna0<-colwise(function(x){
  return(ifelse(is.na(x), 0, x))})  # codifica na=0
# 
# recna10<-colwise(function(x){
#   return(ifelse(is.na(x), 10, x))})  # codifica na=0
# 

#### casos - controles
#casos <- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/casos confirmados 565.xlsx")
casos<-read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/casos confirmados 565 con GMA.xlsx")
names(casos)[c(2,6,7)]<-c("fid","fod","fob")
names(casos)[c(1,4,5,11,12)]<-c("patient_ido","zbs","are","barthel","pgma")  # area no tengo de controles
casos$sex<-ifelse(casos$DES_SEXO=="Masculino",1,2)
casos$pgma<-as.numeric(casos$pgma)
casos<-casos[,-c(3,5,8:10)]  # quito tb area que no lo tengo para controles
## hay una tabla cas con tsi
TSIcas <- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/TSI_CASOS_Casos_a_excluir_deControles.xlsx")
names(TSIcas)[1]<-"patient_ido"
casos<-casos %>% full_join(select(TSIcas,c(1,11)))
names(casos)[9]<-"TIS"
casos$caco<-"ca"



### hitos casos

### inicio y fin del reclutamiento/intervención
summary(casos$fid) #2014-10-03/2015-09-30
### ### inicio y fin de fallecimiento 
casos %>% select(fod) %>% filter(year(fod)<2020) %>% summary(fod)
#2014-10-30/2016-12-04



########################################3 controles

#assign("last.warning", NULL, envir = baseenv())
#controles <- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Posibles controles.xlsx")
controles<- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Controles_TSI_GMA_FECHAS.xlsx")
names(controles)
names(controles)[c(8,2)]<-c("fod","fob")
names(controles)[c(1,6,11:12)]<-c("patient_ido","zbs","pgma","gma5")  # mirar si esto del peso gma es peso.. y en casos si tendremos algo parecido a perfil -entiendo perfilgma
controles$sex<-ifelse(controles$DES_SEXO=="Masculino",1,2)
names(controles)[9]<-"TIS"
controles<-controles[,-c(3:5,7,10,13)] # cosubfar?



### hitos controles

### ### inicio y fin de fallecimiento 
controles %>% select(fod) %>% filter(year(fod)<2020) %>% summary(fod)
#2014-04-10/2017-05-25
#patcon %>% select(fecha) %>% summary(fecha) 
#1800-12-28 /2017-05-27 


# añado barthel
barcon<- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Barthel_controles_eval2y.xlsx")
barcon$FECHA<- as.Date(as.character(barcon$FECHA), "%Y%m%d")
barcon<-select(barcon,-2)
names(barcon)<-c("patient_ido","fechab","barthel")
barcon<-barcon %>% arrange(patient_ido, desc(fechab)) %>% filter(!duplicated(patient_ido))  # nos quedamos con el último..
# barthel controles se tiene para 4976/15248
controles<-left_join(controles,barcon[,-2])
controles<-anti_join(controles,casos,by = "patient_ido")  # 15020
controles$caco<-"co"




rm(barcas,barcon)
rm(TSIcas)



# 
# Patologías	CIAP 2
# Demencia	'P70'
# Diabetes	'T90','T89'
# Insuficiencia cardiaca	'K77'
# Cardiopatía isquémica	K74','K75','K76'
# Ictus	K90',,'K91',
# Insufiuciencia renal 	'U9901'
# Cirrosis	'D97'
# Asma	R95
# EPOC	R96

## create table of codes
codpat<-data.frame()
COD_CIAP<-c("P70","T89","T90","K77","K74","K75","K76","K90","K91","U9901","D97","R95","R96")
pat9<-c("De","Di","Di","Ic","Ci","Ci","Ci","Av","Av","Ir","Ci","As","Ep")
codpat<-as.data.frame(cbind(COD_CIAP,pat9))


####################################  patcas

### comorbilidad
patcas<-read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/patologias casos.xlsx")
patcas$FECHA_INICIO<- as.Date(as.character(patcas$FECHA_INICIO), "%Y%m%d")

patcas<-left_join(patcas,codpat)
patcas<-patcas %>% filter(!is.na(pat9))  # tb estaban dislipemias

# prescindo de las patologias cronicas posteriores a la fecha de entrada (fid)
names(patcas)[1]<-"patient_ido"
patcas<-left_join(patcas,select(casos,1:2))
#patcas$fid<-as.Date(patcas$fid)
patcas<-filter(patcas,FECHA_INICIO<fid)
## ahora se que todas las cronicas son previas.. paso a formato ancho

patcas<-dcast(patcas, patient_ido ~ pat9)  # de los 565 casos/ 520 tenían alguna de estas patologías
## hay que recodificar, que sea 0/1
## al unirlo con los casos, habrá unos cuantos con NA, pasarlo a 0

patcas[,2:9]<-rec01(patcas[,2:9])

patcas<-left_join(casos[,1],patcas)
patcas[,2:9]<-recna0(patcas[,2:9])


write.table(patcas, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_patcas.txt", sep="\t")



####################################  patcon


patcon<-read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/patologias controles.xlsx")
patcon$FECHA_INICIO<- as.Date(as.character(patcon$FECHA_INICIO), "%Y%m%d")

patcon<-left_join(patcon,codpat)
patcon<-patcon %>% filter(!is.na(pat9))  # tb estaban dislipemias

# aqui hay fechas de las patologias cronicas y se tiene que enlazar con la fid de los casos!!
names(patcon)[1]<-"patient_ido"
patcon<-left_join(patcon,select(controles,1))
names(patcon)[4]<-"fecha"

## podria quitarme todas las patologias de controles posteriores al último caso.. (de hecho el 3er quartil está por debajo del mometo de reclutamiento del primer caso)
## elimino las patcon que aparecen despues del ultimo caso
patcon<-filter(patcon, fecha<"2015-09-30")

patcon %>% select(fecha) %>% mutate(pre=ifelse(fecha<"2014-10-03",1,0)) %>% summary() #  Mean   :0.9271  
## solo queda 7,3% de los procesos cronicos de conbtroles que se diagnostican mientras se diagnostican en el año de reclutamiento de los casos: caben dos opciones.
## una es -dado que no nos importanta la fecha si es anterior al reclutamiento global- recodificar todas las fechas a ese momento inicial.. y se podria montar un dcast con esta configuración [si se usara sólo está prescindiriamos del 7,43% de los diagnósticos]
## para incorporar esto diagnosticos puedo -quizas con una granularidad de mes- incorporar nuevas configuraciones para aquellos que los tuvieron..de modo que si se tuviera que matchear por esta mezcla de dianósticos secundarios-fecha.. ya tendria más oportunidades uno que incorpora nuevo diangnosticos..

##    VOY A CONSIDERAR QUE TODAS ESAS FECHAS SON ANTERIORES A LOS RECLUTAMIENTOS.. CREO UNA TABLA DE COMORBILIDAD DE LOS CONTROLES SIN TENERLA EN CUENTA
patcon<-dcast(patcon, patient_ido ~ pat9) 

# recodificar, que sea 0/1
## al unirlo con los casos, habrá unos cuantos con NA, pasarlo a 0

patcon[,2:9]<-rec01(patcon[,2:9])
patcon<-left_join(controles[,1],patcon)
patcon[,2:9]<-recna0(patcon[,2:9])

write.table(patcon, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_patcon.txt", sep="\t")



################################
casos<-left_join(casos,patcas)
controles<-left_join(controles,patcon)

rm(patcas,patcon)
rm(COD_CIAP,pat9,codpat)
# 
# ############################## perfil gma
# http://dwoll.de/rexrepos/posts/recode.html
# GMA 5	Peso> 19,999
# GMA 4	13,077>Peso<19,1000
# GMA 3	9,973>Peso<13,077
# GMA 2	3,062>Peso<9,973
# GMA 1	Peso<3,062



casos$gma5<-cut(casos$pgma, breaks =c(3.062, 9.973, 13.077, 19.1, 19.999, 50))
controles$gma5<-cut(controles$pgma, breaks =c(3.062, 9.973, 13.077, 19.1, 19.999, 50))

### nueva información: dependencia
## nuevo criterio para excluir casos: sólo pluripatológicos







```

```{r nav_ ingresos, urg, vis, com}



#https://www.r-bloggers.com/aggregation-with-dplyr-summarise-and-summarise_each/
#http://marcoghislanzoni.com/blog/2014/09/01/pivot-tables-r-dplyr/

# 
# ** con todas estas medidas de ingresos/consumo de fármacos.. para poder hacerlos comparables caco sin que sean tiempo-dependientes.. prescindimos del periodo de reclutamiento.. se elimina todo lo posterior a esa fecha (#2014-10-03)

##### ingresos hospitalarios  ###################################################################################################
################################## controles
Inghcon <- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Ingresos_controles_eval2y.xlsx")
Inghcon$FECHA_INGRESO<- as.Date(as.character(Inghcon$FECHA_INGRESO), "%Y%m%d")
Inghcon$FECHA_ALTA<- as.Date(as.character(Inghcon$FECHA_ALTA), "%Y%m%d")
names(Inghcon)<-c("patient_ido","fing","falt","com","ting","thos","te")
Inghcon<-filter(Inghcon, fing<"2014-10-03")
# nos quedamos sólo con los ingresos urgentes y hos
Inghcon<-filter(Inghcon, ting=="URG")  ## en teoria debia ser HPE!!
Inghcon<-filter(Inghcon, thos=="H") 
#table(Inghcon$thos) #  D    H 
bypat<-group_by(Inghcon,patient_ido)
#nhco<-dcast(i, patient_ido ~ pat9
nhco<-dplyr::summarise(bypat, nhu=n())
write.table(Inghcon[,-c(4)], "~/Dropbox/Proyectos/cronicos_nav/datos/nav_inghcon.txt", sep="\t")

#controles<-controles[,-(20:21)]
controles<-left_join(controles,nhco)
controles[,19]<-recna0(controles[,19])


################################## casos

Inghcas <- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Ingresos_casos_eval2y.xlsx")
Inghcas$FECHA_INGRESO<- as.Date(as.character(Inghcas$FECHA_INGRESO), "%Y%m%d")
Inghcas$FECHA_ALTA<- as.Date(as.character(Inghcas$FECHA_ALTA), "%Y%m%d")
names(Inghcas)<-c("patient_ido","fing","falt","com","ting","thos","te")
Inghcas<-filter(Inghcas, fing<"2014-10-03")
# nos quedamos sólo con los ingresos urgentes y hos
Inghcas<-filter(Inghcas, ting=="URG", thos=="H")  ## en teoria debia ser HPE!!
#table(Inghcon$thos) #  D    H 
bypat<-group_by(Inghcas,patient_ido)
nhca<-dplyr::summarise(bypat, nhu=n())
write.table(Inghcas[,-c(4)], "~/Dropbox/Proyectos/cronicos_nav/datos/nav_inghcas.txt", sep="\t")

#casos<-casos[,-21]
casos<-left_join(casos,nhca)
casos[,20]<-recna0(casos[,20])

rm(Inghcas,Inghcon,bypat,nhca,nhco)



#### urgencias #####################################################################################################333
urgcon <- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Urgencias_controles_eval2y.xlsx")
urgcon$FECHA<- as.Date(as.character(urgcon$FECHA), "%Y%m%d")
names(urgcon)[1:2]<-c("patient_ido","fvisu")
urgcon<-filter(urgcon, fvisu<"2014-10-03")
bypat<-group_by(urgcon,patient_ido)
#nhco<-dcast(i, patient_ido ~ pat9
nuco<-dplyr::summarise(bypat, nvu=n())
write.table(urgcon[,-3], "~/Dropbox/Proyectos/cronicos_nav/datos/nav_urgcon.txt", sep="\t")

#controles<-controles[,-21]
controles<-left_join(controles,nuco)
controles[,20]<-recna0(controles[,20])  ## obviamente coincide ingreso urgente con visita a urgencia.. habria que restarlas!!
controles$nvu<-controles$nvu-controles$nhu
controles[,20]<-recn0(controles[,20])

#table(controles$nvu)

############3 

urgcas <- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Urgencias_casos_eval2y.xlsx")
urgcas$FECHA<- as.Date(as.character(urgcas$FECHA), "%Y%m%d")

names(urgcas)[1:2]<-c("patient_ido","fvisu")
urgcas<-filter(urgcas, fvisu<"2014-10-03")
bypat<-group_by(urgcas,patient_ido)
#nhco<-dcast(i, patient_ido ~ pat9
nuca<-dplyr::summarise(bypat, nvu=n())

write.table(urgcas[,-3], "~/Dropbox/Proyectos/cronicos_nav/datos/nav_urgcas.txt", sep="\t")


casos<-left_join(casos,nuca)
casos[,21]<-recna0(casos[,21])
casos$nvu<-casos$nvu-casos$nhu
casos[,21]<-recn0(casos[,21])

table(casos$nvu)
### voy a dejar en nvu las visitas que no acabaron en ingreso!! :-( me salen negativos -lo que quiere decir que tuvo ingresos urgentes que no computaron visitas a urgencia (llegaria x ambulancia...)

rm(urgcas, urgcon,bypat,nuca,nuco)
#rm(nhco)


### medicamentos #####################################################################################################333
Medcont<- read_delim("~/Dropbox/Proyectos/cronicos_nav/datos/Medic_disp_controlescronic.csv",  ";", escape_double = FALSE, trim_ws = TRUE)

Medcont<-select(Medcont,c(1,4,5))
names(Medcont)<-c("CIPNA_ANONIMO","ATC","FECHA")
Medcont$FECHA<- as.Date(as.character(Medcont$FECHA), "%Y%m%d")
names(Medcont)<-c("patient_ido","ATC","ffa")

## primero filtro por fecha.. elimino lo anterior al programa
medcon<-filter(Medcont, ffa<"2014-10-03")
medcon$y<-year(medcon$ffa)
medcon$ATC3<-substr(medcon$ATC,1,3)
rm(Medcont)

write.table(medcon, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_medcon.txt", sep="\t")

#rm(dmedcon)

## nos quedamos con esos 3 primeros meses de 2014, diferentes tipos de farmacos (atc3) y numero de ellos 

cmedcon<-dcast(medcon, patient_ido + ATC3~ y)
  cmedcon<-cmedcon[,c(1,2,6)]
 names(cmedcon)[3]<-"y04" 
byatc3<-group_by(cmedcon,patient_ido)
atc3a4<-dplyr::summarise(byatc3, tatc3a4=n(), sum4=sum(y04))


controles<-left_join(controles,atc3a4)
controles[,21:22]<-recna0(controles[,21:22])  


#rm(Inghcon)
rm(atc3a4,cmedcon)
rm(byatc3)
rm(medcon)


############################### casos

Medcas <- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Medic_disp_casos.xlsx")
Medcas<-select(Medcas,c(1,4,5))
names(Medcas)<-c("patient_ido","ATC","ffa")
Medcas$ffa<- as.Date(as.character(Medcas$ffa), "%Y%m%d")


## primero filtro por fecha.. elimino lo anterior al programa
medcas<-filter(Medcas, ffa<"2014-10-03")
rm(Medcas)
medcas$y<-year(medcas$ffa)
medcas$ATC3<-substr(medcas$ATC,1,3)

write.table(medcas, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_medcas.txt", sep="\t")

## nos quedamos con esos 3 primeros meses de 2014, diferentes tipos de farmacos (atc3) y numero de ellos 

cmedcas<-dcast(medcas, patient_ido + ATC3~ y)
  cmedcas<-cmedcas[,c(1,2,6)]
 names(cmedcas)[3]<-"y04" 
byatc3<-group_by(cmedcas,patient_ido)
atc3a4<-dplyr::summarise(byatc3, tatc3a4=n(), sum4=sum(y04))


casos<-left_join(casos,atc3a4)
casos[,22:23]<-recna0(casos[,22:23])  


rm(atc3a4,cmedcas,byatc3,medcas)




### consultas a especializada ###################################################################################

############################################ controles

conescon <- read_delim("~/Dropbox/Proyectos/cronicos_nav/datos/Controles_consultas_med_enf_AE.csv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)
conescon$FECHA<- as.Date(as.character(conescon$PK_FECHA_CITA), "%Y%m%d")
conescon<-select(conescon,c(1,3,7))
names(conescon)<-c("patient_ido","tve","fve")
conescon<-filter(conescon,tve=="MEDICA")  # me quedo, para el matching-- solo con las medicas

## primero filtro por fecha.. elimino lo anterior al programa
conescon<-filter(conescon, fve<"2014-10-03")
cescon<-dcast(conescon, patient_ido ~ tve)
names(cescon)[2]<-"vme"
#escas<-mutate(cescas,tve=vehd+vee+vme)


write.table(cescon, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_conescon.txt", sep="\t")


controles<-left_join(controles,cescon)
controles[,23]<-recna0(controles[,23])  

rm(cescon,conescon)


############################################ casos

conescas <- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Casos_Consultas_med_enf_AE.xlsx")
conescas$FECHA<- as.Date(as.character(conescas$PK_FECHA_CITA), "%Y%m%d")
conescas<-select(conescas,c(1,3,7))
names(conescas)<-c("patient_ido","tve","fve")
conescas<-filter(conescas,tve=="MEDICA")  # me quedo, para el matching-- solo con las medicas

## primero filtro por fecha.. elimino lo anterior al programa
conescas<-filter(conescas, fve<"2014-10-03")
cescas<-dcast(conescas, patient_ido ~ tve)
names(cescas)[2]<-c("vme")
#escas<-mutate(cescas,tve=vehd+vee+vme)

write.table(cescas, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_conescas.txt", sep="\t")

casos<-left_join(casos,cescas)
casos[,24]<-recna0(casos[,24])  


rm(cescas,conescas)



### consultas a primaria ##############################################################3
###################################################3333  casos
conapcas <- read_delim("~/Dropbox/Proyectos/cronicos_nav/datos/ConsultasAP_casos_evalcronic.csv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)
conapcas$Fecha<- as.Date(as.character(conapcas$Fecha), "%Y%m%d")
conapcas<-select(conapcas,c(1:2))
names(conapcas)<-c("patient_ido","fvp")
## primero filtro por fecha.. elimino lo anterior al programa
conapcas<-filter(conapcas, fvp<"2014-10-03")

write.table(conapcas, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_conapcas.txt", sep="\t")

## me quedo con el 2004, 3 meses
conapcas$y<-year(conapcas$fvp)
conapcas<-filter(conapcas,y==2014)
conapcas<-dcast(conapcas, patient_ido ~ y)
names(conapcas)[2]<-c("vap")  # visitas a ap en 2014

#names(casos)[26]<-"vap"
casos<-left_join(casos,conapcas)
casos[,25]<-recna0(casos[,25])  

rm(conapcas)

###################################################3333  controles
conapcon <- read_delim("~/Dropbox/Proyectos/cronicos_nav/datos/ConsultasAP_controles_evalcronic.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE)
conapcon<-select(conapcon,1:2)
names(conapcon)<-c("patient_ido","fvp")
conapcon$Fecha<- as.Date(as.character(conapcon$fvp), "%Y%m%d")
conapcon<-conapcon[,-2]
names(conapcon)<-c("patient_ido","fvp")

write.table(conapcon, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_conapcon.txt", sep="\t")

conapcon$y<-year(conapcon$fvp)
conapcon<-filter(conapcon,y==2014)
conapcon<-dcast(conapcon, patient_ido ~ y)
names(conapcon)[2]<-c("vap")  # visitas a ap en 2014

#names(casos)[26]<-"vap"
controles<-left_join(controles,conapcon)
controles[,24]<-recna0(controles[,24])  

rm(conapcon)

#codid<-codid[,c(1,4)]

# codid<-bind_rows(casos[,c(1,10)], controles[,c(1,9)])
# codid$patid<-as.character(1:dim(codid)[1])
# controles<-left_join(controles,codid, by="patient_ido")
# casos<-left_join(casos,codid)


#####################################################33 dependencia  // prescindo de la fecha!!

conapcas$Fecha<- as.Date(as.character(conapcas$Fecha), "%Y%m%d")

####################################  depcon  
#newDateTime <- as.Date(helpData[ , "ExcelNum"], origin = "1899-12-30")

depcon<-read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Dependencia_controles.xlsx")
names(depcon)<-c("patient_ido","sitdep","fecha")
depcon$fecha<-as.Date(depcon$fecha, origin = "1899-12-30")
depcon$sitdep<-as.factor(depcon$sitdep)

controles<-left_join(controles,select(depcon,1:2))
levels(controles$sitdep)<-c(levels(controles$sitdep),"No valorado")
controles$sitdep[is.na(controles$sitdep)]<-"No valorado"


write.table(depcon, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_depcon.txt", sep="\t")

rm(depcon)
####################################  depcas

depcas<-read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Dependencia_casos.xlsx")
names(depcas)<-c("patient_ido","sitdep","fecha")
depcas$fecha<-as.Date(depcas$fecha, origin = "1899-12-30")

depcas$sitdep<-as.factor(depcas$sitdep)
casos<-left_join(casos,select(depcas,1:2))
levels(casos$sitdep)<-c(levels(casos$sitdep),"No valorado")
casos$sitdep[is.na(casos$sitdep)]<-"No valorado"


write.table(depcas, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_depcas.txt", sep="\t")
rm(depcas)



```



```{r matching dataset con dependencia, y exclusión de controles no pluripatológicos en dat -la base que auna casos y controles}


############33 controles a excluir
ezcon<- read_excel("~/Dropbox/Proyectos/cronicos_nav/datos/Cipnas_excluir_controles.xlsx")


ezcon<-ezcon %>% select(CIPNA_ANONIMO)
names(ezcon)<-"patient_ido"
controles<-anti_join(controles,ezcon)  # 14368

rm(ezcon)




write.table(casos, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_casos.txt", sep="\t")
write.table(controles, "~/Dropbox/Proyectos/cronicos_nav/datos/nav_controles.txt", sep="\t")


save.image("caco.Rdata")
save(namex, file="namex")

load("~/Dropbox/Proyectos/cronicos_nav/caco.Rdata")
load("namex")


#####################################################3 dat: union de casos y controles
dat<-rbind(casos[,-c(2)],controles) # 14933


dat$yb<-lubridate::year(dat$fob)
dat$zbs<-as.factor(dat$zbs)
dat$sex<-as.factor(dat$sex)
dat$TIS<-as.factor(dat$TIS)


dat$CC<-as.numeric(as.factor(dat$caco))
dat$CC<-ifelse(dat$CC==2,0,1)
rownames(dat)<-dat$patient_ido

### ¿que hacemos con barthel y con gma5, con bastantes missing)
 #    barthel                gma5           TIS       
 # Min.   :  0.00   (3.06,9.97]:3219   TSI 001:  540  
 # 1st Qu.: 35.00   (9.97,13.1]:3217   TSI 002:13247  
 # Median : 60.00   (13.1,19.1]:5317   TSI 003: 1019  
 # Mean   : 59.12   (19.1,20]  : 577   TSI 004:  510  
 # 3rd Qu.: 90.00   (20,50]    :3128   TSI 005:   52  
 # Max.   :100.00   NA's       : 127   TSI 006:  107  
 # NA's   :10297                       NA's   :  110  

## barthel descartado, pero gma5 -110 missing en casos- podria llevarnos a hacer 2 matching con y sin gma5 en el poesc

#summary(dplyr::select(dat,barthel,gma5,TIS))

summary(dat)

```


```{r imputacion}


# no puede haber NA en la base
### le voy a colocar el valor medio de pmga a los na

https://www.analyticsvidhya.com/blog/2016/03/tutorial-powerful-packages-imputing-missing-values/

library(mice)

dat$vivo<-lubridate::year(dat$fod)>2017

dat
summary(dat)

datc3<-select(dat,-zbs,-barthel,-patient_ido,-caco,-gma5,-fod,-fob)
 md.pattern(datc3)  # describe patrones de missing
 
 
 str(datc3)
  imputed_Data1 <- mice(datc3, m=1, maxit = 50, seed = 500)
datc3 <- complete(imputed_Data1)


# library(VIM)
# mice_plot <- aggr(dat, col=c('navyblue','yellow'),
#                     numbers=TRUE, sortVars=TRUE,
#                     labels=names(dat), cex.axis=.7,
#                     gap=3, ylab=c("Missing data","Pattern"))

```



```{r matching}

foi

############################## formula
# 

datc3<-cbind(dat[,1],datc3)

names(datc3)
# 
 pp <-paste(names(datc3)[-c(1,21:22)],"+",collapse="")
namex <-  substr(pp,1,nchar(pp)-1)
 namex

foi <-as.formula(paste("CC ~ ",namex))
foi
# 
# glm1 <- glm(foi, family = binomial, data = dat)
# rm(glm1)
# MatchBalance(foi,data=data)
# 
# 
foi


library(MatchIt)

matchitcc <- matchit(foi, data = datc3, method="nearest", ratio=3)
b <- summary(matchitcc)
b

foi

kable(b$nn, digits = 2, align = 'c', 
      caption = 'Table 2: Sample sizes')


kable(b$sum.matched[c(1,2,4)], digits = 2, align = 'c', 
      caption = 'Table 3: Summary of balance for matched data')


plot(matchitcc, type = 'jitter', interactive = FALSE)




# Saving the matched samples
dfmatch <- match.data(matchitcc)[1:ncol(datc3)]

xx<-dfmatch

xx<-left_join(dfmatch,dplyr::select(dat,patient_ido,fod,gma5,zbs,barthel))


table(xx$vivo, xx$CC)



write.xlsx(xx, "matchcc.xlsx")


rm(xx,datref)
rm(b,imputed_Data1,mice_plot,namex,pp)
save.image("match.Rdata")

```



```{r}




# Matching cases to controls is an important pre-analysis step. The rEHR package provides three methods for matching cases to controls:
# 
# Incidence density matching (IDM)
# Exact matching
# Matching on a dummy index date sourced from consultation files
# 
# 4.1.1 Incidence density matching
# This is performed using the get_matches() function. With IDM, controls are selected for a particular case at the time of diagnosis (or other event such as death) from other members of the cohort who, at that time, do not have the diagnosis. The IDM sampling procedure allows the same patient to be selected as a control for more than one case, thus providing a full set controls for each case while still producing unbiased estimates of risk [7; 21]. This also means that the matching procedure can be parallelised to increase computational efficiency.

# 
# cohort2 <- build_cohort(prevalence_dat, cohort_type =  "incid",
# cohort_start =  "2006‐01‐01",
# cohort_end =  "2012‐12‐31",
# diagnosis_start =  "eventdate")



# 
# names(casos)[1]<-"patido"
# names(controles)[1]<-"patido"
# dim(casos)[1]


casos$fid<-as.character(casos$fid)

names(controles)[3]<-"sex"

controles$fid<-as.Date(controles$FECHA_FALLECIMIENTO, "%Y-%m-%d")
controles$fid<-as.character(controles$fid)  
casos$sex<-as.character(casos$sex)
controles$sex<-as.character(controles$sex)

names(casos)
controles$pp="n"
casos$pp="n"

## aqui hay un tema, y es que si se quieren controles y casos para hacerles una entrevista igual hay que incluir la condicción "no fallecidos" (el estado está definido hace un mes)

IDM_controls <- get_matches(cases =  casos[,c(2,11:14)],
control_pool =  controles[,c(2:3,9:13)],
match_vars =  c("sex","pp"),
n_controls =  4, cores =  1,
method =  "incidence_density",
diagnosis_date =  "fid")

IDM_controls

IDM_controls$matched_case


head(casos)
c

??get_matches
??match_case

match_case()


IDM_controls <- get_matches(cases =  filter(cohort2, case == 1),
control_pool =  filter(cohort2, case == 0),
match_vars =  c("gender", "region"),
n_controls =  4, cores =  1,
method =  "incidence_density",
diagnosis_date =  "eventdate")

```




```{r sqlite}


nav_path <- "~/Dropbox/Proyectos/cronicos_nav/datos"


## create a new database connection to a temporary file
db <- database(tempfile(fileext = ".sqlite"))
## Import multiple data files into the database
import_CPRD_data(db, data_dir = nav_path, filetypes = c("casos","controles","inghcon","inghcas","urgcon","urgcas","patcas","patcon","conescas","conescon","medcon","medcas","barcon","barcas","conapcon","conapcas"), dateformat = "%Y-%m-%d", yob_origin = 1800, regex = "nav", recursive = TRUE)

rm(casos)

#practices <- select_events(db = db, tab = "Practice", convert_dates = TRUE)
casos<-select_events(db, tab = "casos",convert_dates = TRUE)

controles<-select_events(db, tab = "controles")


#barcon<-select_events(db, tab = "barcon",convert_dates = TRUE)




### de controles hay que excluir los casos.. y tb aquellos que provienen de algunas áreas...
#casos %>% mutate(sexo=mapvalues(DES_SEXO,c("Femenino","Masculino"),c(1,2)))


```

```{r propensity score}

http://sekhon.berkeley.edu/papers/MatchingJSS.pdf
https://datascienceplus.com/how-to-use-r-for-matching-samples-propensity-score/


library("Matching")
data("lalonde")
 attach(lalonde)
Y <- lalonde$re78
Tr <- lalonde$treat
#We now estimate our first propensity score model:
glm1 <- glm(Tr ~ age + educ + black + hisp + married + nodegr + re74 + re75, family = binomial, data = lalonde)
#Let us do one-to-one matching with replacement using our preliminary propensity score model where the estimand is the average treatment effect on the treated (ATT):
rr1 <- Match(Y = Y, Tr = Tr, X = glm1$fitted)
summary(rr1)

 m1 = Match(Y = Y, Tr = Tr, X = glm1$fitted,  ties=FALSE)
 
MatchBalance(Tr ~ age + I(age^2) + educ + I(educ^2) + black + hisp + married + nodegr + re74 + I(re74^2) + re75 + I(re75^2) + u74 + u75 + I(re74 * re75) + I(age * nodegr) + I(educ * re74) + I(educ * re75), match.out = rr1, nboots = 1000, data = lalonde)



 m1$index.treated
 m1$index.control
 
 library(wakefield)  ## permite crear una base simulada
 
 ## pacientes
 set.seed(1234)
df.patients <- r_data_frame(n = 250, 
                            age(x = 30:78, 
                                name = 'Age'), 
                            sex(x = c("Male", "Female"), 
                                prob = c(0.70, 0.30), 
                                name = "Sex"))
df.patients$Sample <- as.factor('Patients')
 
## poblacion
set.seed(1234)
df.population <- r_data_frame(n = 1000, 
                              age(x = 18:80, 
                                  name = 'Age'), 
                              sex(x = c("Male", "Female"), 
                                  prob = c(0.50, 0.50), 
                                  name = "Sex"))
df.population$Sample <- as.factor('Population')


#Merging the dataframes

 # Based on the variable Sample, we create a new variable named Group (type logic) and a further variable (Distress) containing information about the individuals’ level of distress. The Distress variable is created using the age-function of the Wakefield package. As we can see, women will have higher levels of distress.

mydata <- rbind(df.patients, df.population)
mydata$Group <- as.logical(mydata$Sample == 'Patients')
mydata$Distress <- ifelse(mydata$Sex == 'Male', age(nrow(mydata), x = 0:42, name = 'Distress'),
                                                age(nrow(mydata), x = 15:42, name = 'Distress'))
library(tableone)
table1 <- CreateTableOne(vars = c('Age', 'Sex', 'Distress'), 
                         data = mydata, 
                         factorVars = 'Sex', 
                         strata = 'Sample')
table1 <- print(table1, 
                printToggle = FALSE, 
                noSpaces = TRUE)

library(knitr)
kable(table1[,1:3],  
      align = 'c', 
      caption = 'Table 1: Comparison of unmatched samples')


#Matching the samples ###############3

# Now, that we have completed preparation and inspection of data, we are going to match the two samples using the matchit-function of the MatchIt package. The method command method="nearest" specifies that the nearest neighbors method will be used. Other matching methods are exact matching, subclassification, optimal matching, genetic matching, and full matching (method = c("exact", "subclass", "optimal", ""genetic", "full")). The ratio command ratio = 1 indicates a one-to-one matching approach. With regard to our example, for each case in the patient sample exactly one case in the population sample will be matched. Please also note that the Group variable needs to be logic (TRUE vs. FALSE).
 
library(MatchIt)
set.seed(1234)
match.it <- matchit(Group ~ Age + Sex, data = mydata, method="nearest", ratio=1)
a <- summary(match.it)


kable(b$sum.matched[c(1,2,4)], digits = 2, align = 'c', 
      caption = 'Table 3: Summary of balance for matched data')


names(a$sum.matched)


plot(match.it, type = 'jitter', interactive = FALSE)
plot(matchitcc, type = 'jitter', interactive = FALSE)


# Saving the matched samples
df.match <- match.data(match.it)[1:ncol(mydata)]
rm(df.patients, df.population)


##----------------------------------------------------------->>>>
dfmatchcc<-match.data(matchitcc)[1:ncol(data)]

table4 <- CreateTableOne(vars = c('Age', 'Sex', 'Distress'), 
                         data = df.match, 
                         factorVars = 'Sex', 
                         strata = 'Sample')
table4 <- print(table4, 
                printToggle = FALSE, 
                noSpaces = TRUE)
kable(table4[,1:3],  
      align = 'c', 
      caption = 'Table 4: Comparison of matched samples')


 ```




```{r Building cohorts, matching and time-varying covariates}

# hay que caracterizar a casos y controles, 

# a) agregando las variables de conteo necesarias para el matching: 
# -	Nº ingresos y días de estancia de HPE 
  ## ->inghcas/con -> extraer para los casos los ingresos previos a la FECHA_ALTA_ES_CRONICO: 
        # --> ¿y como hacemos para los controles.. tendria que añadir en inghcon unn campo con el acumulado de ingresos!
        # y con los dias de estancia ¿sacamos también el acumulado para casos y la frecuencia acumulada segun fecha para controles?
# -	Nº consultas de medicina en AE
# -	Nº consultas de enfermería en AE
# -	Nº fármacos crónicos <- ¿entiendo que hablamos de principio ATC.. puedo sumar los diferentes tipos de la variable ATC?


practices <- select_events(db = db, tab = "Practice", convert_dates = TRUE)
prevalence_dat <- left_join(prevalence_dat, practices)
cohort <- build_cohort(prevalence_dat, cohort_type = "prev", cohort_start = "2006-01-01", cohort_end = "2012-12-31", diagnosis_start = "eventdate")


##Error: objeto 'prev_num' no encontrado

consultation_dir <- "~/R/rEHR_testing"
flat_files(db, out_dir = consultation_dir, file_type = "csv")
index_controls <- match_on_index(cases = filter(cohort2, case == 1),control_pool = filter(cohort2, case == 0),index_var = "eventdate",match_vars = c("gender", "region"),index_diff_limit = 90,consult_path = consultation_dir,n_controls = 4,import_fn = function(x) convert_dates(read.csv(x)))unlink(consultation_dir, recursive = TRUE) # clean up constructed dirs after analysis


??match_case

match_case()

```

```{r cut_tv}

tv_test <- data.frame(id = 1:5, start = rep(0, 5), end = c(1000, 689, 1000, 874, 777), event = c(0,1,0,1,1), drug_1 = c(NA, NA, NA, 340, 460), drug_2 = c(NA, 234, 554, 123, NA), drug_3_start = c(110, 110,111, 109, 110), drug_3_stop = c(400, 400, 400, 400, 400),stage_1 = c(300, NA, NA, NA, NA),stage_2 = c(450, NA, NA, NA, NA))

## Multiple binary chronic covariates:
tv_out1 <- cut_tv(tv_test, entry = start, exit = end, cut_var = drug_1, id_var = id, tv_name = drug_1_state)
tv_out1 <- cut_tv(tv_out1, start, end, drug_2, id_var = id, drug_2_state)

head(tv_out1)

# Binary covariates:
tv_out3 <- cut_tv(tv_test, start, end, drug_3_start, id_var = id, drug_3_state)
tv_out3 <- cut_tv(tv_out3, start, end, drug_3_stop, id_var = id, drug_3_state)
head(tv_out3)

## incremental covariates:
inc_1 <- cut_tv(tv_test, start, end, stage_1, id_var = id, disease_stage, on_existing = "inc")
inc_1 <- cut_tv(inc_1, start, end, stage_2, id_var = id, disease_stage, on_existing = "inc")
head(inc_1)

## Chaining combinations of the above using %>%
library(dplyr)
tv_test %>%
cut_tv(start, end, drug_1, id_var = id, drug_1_state) %>%
cut_tv(start, end, drug_2, id_var = id, drug_2_state) %>%
cut_tv(start, end, drug_3_start, id_var = id, drug_3_state) %>%
cut_tv(start, end, drug_3_stop, id_var = id, drug_3_state) %>%
cut_tv(start, end, stage_1, id_var = id, disease_stage, on_existing = "inc") %>%
cut_tv(start, end, stage_2, id_var = id, disease_stage, on_existing = "inc") %>%
head

```

```{r  Querying the database Selecting all events}
# Once EHR data has been imported to the database, the rEHR package has a number of flexible built-in querying functions for extracting data. These functions are much faster to execute and less error prone than having to loop through hundreds of text files.
# The primary generic query function is select_events() and is able to select all the events in a database table matching a provided where argument. This function is also called by the other more specific query functions. An example set of lists of clinical codes for a number of medical conditions is provided with the package (data(clinical_codes)). select_events() returns a dataframe of extracted data.

diabetes_codes <- clinical_codes[clinical_codes$list == "Diabetes",] 
select_events(db, tab = "Clinical", columns = c("patid", "eventdate", "medcode"), where = "medcode %in% .(diabetes_codes$medcode) & eventdate < '2006-01-01' & eventdate >= '2005-01-01'")

# The user must supply a string representation of valid R code, which is then translated to SQL via
# the dplyr::translate_sql_q function. There are two important caveats to this:
# 
#   1. If an element of the clause represents an R object to be accessed (such as the elements of a vector) it must be wrapped in a .() (See the example above). String elements wrapped in .() are processed by the expand_string function before being passed to dplyr::translate_sql_q.
# ------------>  2. Dates should separately quoted and entered in ISO format (‘%Y-%m-%d’). This is because dates are stored as ISO text in the database, not as r Date types


Asthma_codes <- clinical_codes[clinical_codes$list == "Asthma",]


q <- select_events(db, tab = "Clinical", columns = c("patid", "eventdate", "medcode"), where = "medcode %in% .(Asthma_codes$medcode)", sql_only = TRUE)


temp_table(db, tab_name = "Asthma", select_query = q)
head(db, temp = TRUE)
head(db, table = "Asthma")

# Using raw SQL queries
# Since EHR data is stored as a standard SQLite database, users can alternatively make SQL queries to the
# database using sqldf, which is imported into the namespace on loading of the rEHR package:
# sqldf("SELECT patid, practid, gender, yob, deathdate from Patient WHERE deathdate IS NOT NULL LIMIT 6", connection = db)

# There are two methods for including R objects in raw SQL strings. First, wrapping the string in a call to expand_string() allows for the .() notation to be used as in where arguments to select_events() based functions. Alternatively, a helper function, wrap_sql_query() is provided that functions in a similar way to base::sprintf but formats objects according to SQL syntax. If the result of evaluating the argument is a vector of length 1, it is inserted as is; if it is a vector of length > 1, it is wrapped in parentheses and comma separated.
# 
# medcodes1 <- 1:5
# practice <- 255
# expand_string("SELECT * FROM clinical WHERE practid == .(practice)")
# ## [1] "SELECT * FROM clinical WHERE practid == 255"
# wrap_sql_query("SELECT * FROM clinical WHERE practid == #1 AND medcodes in #2",
# practice, medcodes1)
## [1] "SELECT * FROM clinical WHERE practid == 255 AND medcodes in ( 1, 2, 3, 4, 5 )"

# Selecting first or last events

# Frequently, users need to find the first clinical event for a given patient (e.g. to identify dates of diagnosis of chronic diseases) or the most recent clinical event (e.g. to identify if a drug therapy has been prescribed within a certain time period). rEHR provides convenience functions for these common situations. The functions run a select_events() query and then group by patient id and selects only the earliest/latest event for each patient:

first_DM <- first_events(db, tab = "Clinical", columns = c("patid", "eventdate", "medcode"), where = "medcode %in% .(diabetes_codes$medcode)")
last_DM <- last_events(db, tab = "Clinical", columns = c("patid", "eventdate", "medcode"), where = "medcode %in% .(diabetes_codes$medcode)")
head(first_DM)


# Querying longitudinal data with select_by_year()

# To show the utility of the package we demonstrate how one might extract an incident and prevalent cohort of diabetes patients from the simulated example data. Prevalent events for a chronic condition are selected by the earliest diagnostic event prior to the end of the time period in question. The denominator for the calculation of the prevalence is the total number of patients registered at that time point.

# Select all patients with current registration date (crd) < the start date
# for each year.
registered_patients <- select_by_year(db = db, tables = "patient", columns = c("patid", "practid", "gender", "yob", "crd", "tod", "deathdate"), where = "crd < STARTDATE", year_range = c(2008:2012), year_fn = standard_years)

str(registered_patients)
table(registered_patients$year)

incident_cases <- select_by_year(db = db, tables = c("Clinical", "Referral"), columns = c("patid", "eventdate", "medcode"), where = "medcode %in% .(diabetes_codes$medcode) & eventdate <= ENDDATE", year_range = c(2008:2012), year_fn = standard_years,
selector_fn = first_events)

str(incident_cases)

## Remove duplicates across clinical and referral tables:
incident_cases %>%
group_by(patid, year) %>%
arrange(eventdate) %>%
distinct() %>%
ungroup -> incident_cases
## All patients are kept (equivalent to merge(all.x = TRUE))
prevalence_dat <- left_join(registered_patients, incident_cases)


#########  NO FUNCIONA
# prevalence_dat <- prev_terms(prevalence_dat)
# # 
# # Error in as.Date.default(dat[[column]]) :  do not know how to convert 'dat[[column]]' to class “Date”
# 
# totals <- prev_totals(prevalence_dat)
# totals$prevalence$year_counts

```
